The impact of gas bulk rotation on the Lyman-$\alpha$ line

### 2013-10-23 paper/ms.tex
edits in the discussion section

We start by a quantitative discussion of the line morphology. Here we
follow a classification of $Ly\alpha$ profiles made by
\cite{Kulas12}. They presented observational results for star forming
galaxies at $z=2$-$3$ that showed multiple peaks in their emission.
They classified the observed lines into four groups. In Group I the
red peak is stronger than the blue, in Group II the blue peak is
stronger, in Group III the two peaks have the same strength and Group
IV LAEs present multiple peaks.

Neither Group I nor II  and III can be fully explained by rotation. However,
 some comments about some of the galaxies in this groups must be done, in 
 particular for those galaxies in which the flux in between the two peaks is
 high rotation must be present due to the fact this is until now the most 
 efficient mechanism which produces this effect.

The same argument explained above applies to group IV in which the flux
in between of the two peaks is high ( $ \geq 70 \%$ of the highest peak), 
concerning group V in which multiple-peaks appear homogeneous and central
models can not reproduce such a profiles. While the off-center model could 
generate similar profiles which is the case of (put reference when the fig.
is ready) when tree peaks could appear but with roughly the same flux, this
multiple-peaks still need a better comprehension of the gas kinematics.

An other interesting sample of LAEs was presented by \cite{Yamada2012} 
for 91 LAEs at $z=3.1$. From that sample 31 LAEs ($\sim 33 \%$) present 
a single peak profile, if in this single peak there is an asymmetry it 
can be explained with the outflow/inflow model \cite{Verhamme2008, Dijkstra06}.
 But if it is not the case now they can be understanding with rotation, 
 in particular in this sample there are single peaks with a strong asymmetry 
 and there are peaks that this asymmetry its not notorious.

Between the 91 LAEs 8 reveal profiles with a central peak and two smaller
peaks at the redder and bluer side, this also could be explained by our
off-center model when higher velocities are involved in rotation.

Even in profiles where the double-peaked is presented rotation also could 
be present as it is seen in Fig.\ref{fig:differentvelocities} where rotation
makes that the central part of the line is higher than in profiles where 
there is no rotation. This could be the case of 13 LAEs in the sample of 
\cite{Kulas12} where there is a double-peak profile with similar heigh and 
a central notorious part.

Until now when single peaked profiles are present we associate this to
the presence of inflows and outflows due to the asymmetry of the line
, but in the case in which the profile have symmetry we could say that
gas bulk rotation is the principal effect governing the galaxy. For this reason 
a detailed study involving inflows/outflows and rotation kinematics would 
give as better understanding of the $\mathrm{Ly} \alpha$ profiles we 
measured.

>>>>

We start by a quantitative discussion of the line morphology. Here we
follow a classification of $Ly\alpha$ profiles made by
\cite{Kulas12}. They presented observational results for star forming
galaxies at $z=2$-$3$ that showed multiple peaks in their emission.
They classified the observed lines into four groups. In Group I the
red peak is stronger than the blue, in Group II the blue peak is
stronger, in both cases the peaks are symmetrically located around the
line center. In Group III the blue peak also dominates but there is a
overall redward shift of the zero point. Group IV presents two similar
peaks symmetrically located around the zero point. Group V presents
three peaks.

The asymmetries in Groups I, II  and III cannot be fully explained by
rotation. However, some galaxies in those groups have a high flux between
the two peaks; an effect that can be induced by rotation. The same
argument applies to galaxies in the group IV in which the flux
in between of the two peaks is high ($ \geq 70 \%$ of the intensity in
the highest peak). Finally, the triple peaked lines in Galaxies in
Group V can be easily reproduced by rotation with off-centered
emission.

Another relevant sample of lines was presented by \cite{Yamada2012} 
for 91 LAEs at $z=3.1$. They find that $\sim 33 \%$ of the galaxies present 
a single peak profile. Some of the peaks have a stron symmetry, while
others do not. The symmetric single peaked progile can be explained by
rotation effects, while the assymetric cases naturally arise in
outflow/inflow models \citep{Verhamme2008, Dijkstra06}. Among the 91
LAEs, 8 reveal triple peaked profiles with a central peak and two
smaller peaks at the redder and bluer side, a feature that is present
in our off-center emission models with rotation.

The presence of single peaked profiles has been associated to
inflow/outflow dynamics \citep{XXX}. In this paper we show that gas
bulk rotation can also be considered as a probable origin for that
behaviour, provided that the observed single peak is highly
symmetric. Similarly, in the case of double peaked lines with a high
level of flux at the line center, rotation also deserveds to be considered in
the pool of possible bulk flows responsible for that feature,
specially if the two peaks have similar intensities. The case of
triple peaked lines is another clear feature of gas rotation, this
time with the additional condition of an off-centered emission.
